---
title: "P8105 Homework 1"
author: "Madison Goldrich"
date: "2023-09-14"
output: github_document
---

# Problem 1

## Load library and dataset

Load the `moderndive` library, and load the `early_january_weather` dataset.

```{r}
library(moderndive)

data("early_january_weather")
```

## Short description of the dataset

According to the dataset help file, `early_january_weather` contains hourly meteorological data for LGA, JFK, and EWR.

### Summary of variables

Here are the variables included in the dataset:

* `origin` is the weather station. This is EWR for all data points.
* `year`, `month`, `day`, and `hour` represent the time of the weather recording.
  * `year` is `r mean(early_january_weather$year)`.
  * `month` is January, represented as `r mean(early_january_weather$month)`.
* `temp` and `dewp`are the temperature and dewpoint in F.
* `humid` is relative humidity.
* `wind_dir`, `wind_speed`, and `wind_gust` are wind direction (in degrees), speed and gust speed (in mph).
* `precip` is preciptation, in inches.
  * Mean: `r mean(early_january_weather$precip)` inches
  * Standard deviation: `r sd(early_january_weather$precip)` inches
* `pressure` is sea level pressure in millibars.
* `visib` is visibility in miles. 
  * Mean: `r mean(early_january_weather$visib)` miles
  * Standard deviation: `r sd(early_january_weather$precip)` miles
* `time_hour` is the date and hour of the recording as a `POSIXct` date.

### Size of dataset

The size of the `early_january_weather` dataset is `r nrow(early_january_weather)` rows by `r ncol(early_january_weather)` columns.

### Mean temperature

The mean temperature recorded within the dataset timeframe is `r mean(early_january_weather$temp)` degrees Fahrenheit.

## Scatterplot

### Create the scatterplot

Make a scatterplot of `temp` (y) vs. `time_hour` (x); color points using the `humid` variable (adding `color = ...` inside of `aes` in your `ggplot` code should help). 

```{r}
library(tidyverse)

ggplot(early_january_weather, aes(x = time_hour, y = temp, color = humid)) + geom_point()
```

### Interpretation

From the scatterplot, we can see that the temperature fluctuates throughout each day, but the temperature is generally trending upward over the timeframe. Humidity appears to stay fairly low other than a period of a few days where the humidity is higher. The temperature fluctuates less during that period of higher humidity.

### Save!

Export the scatterplot to the project directory.

```{r}
ggsave("prob1_scatterplot.pdf")
```


# Problem 2

## Create dataframe

Create a dataframe comprised of:

* a random sample of size 10 from a standard Normal distribution
* a logical vector indicating whether elements of the sample are greater than 0
* a character vector of length 10
* a factor vector of length 10, with 3 different factor “levels”

```{r}
prob2_df = tibble(
  vec_norm = rnorm(10),
  vec_logic = c(vec_norm>0),
  vec_char = c("This", "is", "a","character", "vector", "of", "length", "ten", "haha", "yay!"),
  vec_factor = factor(c("small", "medium", "large", "large", "medium", "small", "medium", "large", "small", "small"))
)
```

## Calculate means

Try to take the mean of each variable in your dataframe. What works and what doesn't?

```{r}
mean(pull(prob2_df, var = vec_norm))

mean(pull(prob2_df, var = vec_logic))

mean(pull(prob2_df, var = vec_char))

mean(pull(prob2_df, var = vec_factor))
```

The random sample variable `vec_norm` and the logical variable `vec_logic` work for calculating the mean. The character variable `vec_char` and the factor variable `vec_factor` do not work.

## Convert variables

Write a code chunk that applies the `as.numeric` function to the logical, character, and factor variables (please show this chunk but not the output). What happens, and why? Does this help explain what happens when you try to take the mean?

```{r, eval = FALSE}
as.numeric(pull(prob2_df, var = vec_logic))

as.numeric(pull(prob2_df, var = vec_char))

as.numeric(pull(prob2_df, var = vec_factor))
```

The `as.numeric` function works to convert the logical variable to be a vector of 1s and 0s, which helps explain why the mean can be calculated for a logical variable -- there are corresponding values for `TRUE` and `FALSE`. It does not work to convert the character variable; it returns a vector of NAs and a warning message: `NAs introduced by coercion`. This helps explain why the mean could not be calculated for this variable -- character variables cannot be converted to numeric. Finally, the function did work to convert the factor variable to numbers that correspond to each factor level. However, the mean was not able to be calculated for the factor variable. The `mean` function only works for numeric or logical.